Large-Scale Content-Based Matching of MIDI and Audio Files

### 2015-05-01 paper.tex
Add intro to hashing

The ``cost'' of a DTW path $p, q$ through $S$ is calculated by the sum of the distances between the aligned entries of each sequence, or
$$
\sum_{i = 1}^L S[p[i], q[i]]
$$
As-is, the DTW cost is not an appropriate confidence score for alignment/match quality in our application primarily for two reasons.
First of all, the path length can vary greatly across MIDI/audio file pairs depending on $N$ and $M$.
We therefore compute the \textit{mean} distance between aligned constant-Q spectra across the path instead of the sum of distances:
$$
\frac{1}{L}\sum_{i = 1}^L S[p[i], q[i]]
$$
Secondly, the scale of $S$ (and therefore the DTW cost) depends on how similar the synthesized audio is to the ``real'' audio recording, which is not an important factor in the alignment quality.
This can be mitigated by normalizing the DTW cost by the mean value of the submatrix of $S$ through which the DTW past travels:
$$
\mathcal{N} = \sum_{i = \min(p)}^{\max(p)} \sum_{j = \min(q)}^{\max(q)} S[i, j] 
$$
These normalizations were also discussed in \cite{hu2003polyphonic}.
Combining the above, we get our modified DTW cost $c$:
$$
c = \frac{1}{L \mathcal{N}} \sum_{i = 1}^L S[p[i], q[i]]
$$
In practice, we found these two normalization steps to be of critical importance for obtaining a reliable confidence score.

@@

Our hashing model is based on the siamese network architecture proposed in \cite{masci2014multimodal}.
Given feature vectors from two modalities which are labeled as being similar or dissimilar, this approach attempts to learn a nonlinear mapping from each modality to a common Hamming space such that similar and dissimilar feature vectors are mapped to bitvectors with small and large Hamming distances respectively.
This can be expressed as the loss function 
$$
\mathcal{L} = \frac{1}{|\mathcal{P}|} \sum_{(x, y) \in \mathcal{P}} \| f(x) - g(y) \|_2^2 - \alpha \frac{1}{\mathcal{N}} \sum_{(x, y) \in \mathcal{N}} \max(0, m - \|f(x) - g(y) \|_2)^2
$$

>>>>

The ``cost'' of a DTW path $p, q$ through $S$ is calculated by the sum of the distances between the aligned entries of each sequence, or
$$
\sum_{i = 1}^L S[p[i], q[i]]
$$
As-is, the DTW cost is not an appropriate confidence score for alignment/match quality in our application primarily for two reasons.
First of all, the path length can vary greatly across MIDI/audio file pairs depending on $N$ and $M$.
We therefore compute the \textit{mean} distance between aligned constant-Q spectra across the path instead of the sum of distances:
$$
\frac{1}{L}\sum_{i = 1}^L S[p[i], q[i]]
$$
Secondly, the scale of $S$ (and therefore the DTW cost) depends on how similar the synthesized audio is to the ``real'' audio recording, which is not an important factor in the alignment quality.
This can be mitigated by normalizing the DTW cost by the mean value of the submatrix of $S$ through which the DTW past travels:
$$
\mathcal{B} = \sum_{i = \min(p)}^{\max(p)} \sum_{j = \min(q)}^{\max(q)} S[i, j] 
$$
These normalizations were also discussed in \cite{hu2003polyphonic}.
Combining the above, we get our modified DTW cost $c$:
$$
c = \frac{1}{L \mathcal{B}} \sum_{i = 1}^L S[p[i], q[i]]
$$
In practice, we found these two normalization steps to be of critical importance for obtaining a reliable confidence score.

@@

Our hashing model is based on the Siamese network architecture proposed in \cite{masci2014multimodal}.
This approach considers the situation where we have feature vectors from two modalities, and two sets $\mathcal{P}$ and $\mathcal{N}$ which contain pairs of feature vectors from each modality such that if $(x, y) \in \mathcal{P}$ then $x$ and $y$ are considered ``similar'' and if $(x, y) \in \mathcal{N}$ they are ``dissimilar''.
A nonlinear mapping is then learned from each modality to a common Hamming space such that similar and dissimilar feature vectors are respectively mapped to bitvectors with small and large Hamming distances.
One possible objective function for this technique is
\begin{align*}
\mathcal{L} &= \frac{1}{|\mathcal{P}|} \sum_{(x, y) \in \mathcal{P}} \| f(x) - g(y) \|_2^2\\
& - \frac{\alpha}{|\mathcal{N}|} \sum_{(x, y) \in \mathcal{N}} \max(0, m - \|f(x) - g(y) \|_2)^2
\end{align*}
where $f$ and $g$ are the nonlinear mappings for each modality, $\alpha$ is a regularization parameter which controls how important it is that the dissimilar pairs be mapped far apart, and $m$ is a threshold above which dissimilar pairs are considered sufficiently distant.

Finding the nonlinear mappings $f$ and $g$ involves optimizing their parameters with respect to $\mathcal{L}$.
The exact form of these nonlinearities is a system design choice; in \cite{masci2014multimodal} they are chosen to be multi-layer nonlinear networks.
In the present work, we will use convolutional networks due to their ability to exploit invariances in the input feature representation.
CQTs contain invariances in both the time and frequency axes, so convolutional networks are particularly well-suited for our task.

In order to apply this technique, we also need to obtain the collections of pairs similar and dissimilar feature vectors $\mathcal{P}$ and $\mathcal{N}$.
In our problem, our two feature modalities are CQTs from synthesized MIDI files and CQTs from audio files.
A logical choice for $\mathcal{P}$ is to choose pairs of constant-Q spectra from simultaneous beats in successfully aligned MIDI/audio pairs.
For example, if $A, B \in \mathbb{R}^{N \times D}$ are CQTs of an aligned MIDI and audio file respectively, then $\mathcal{P}$ will include $(A[1], B[1]), \ldots, (A[N], B[N])$.
The choice of $\mathcal{N}$ is less obvious.
We found that simply randomly choosing CQT spectra from non-aligned beats in our collection of aligned MIDI and audio files achieved satisfactory results.
